---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# rgsrs

<!-- badges: start -->
[![R-CMD-check](https://github.com/c1au6i0/rgsrs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/c1au6i0/rgsrs/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Tidy R client for the [FDA Global Substance Registration System (GSRS)](https://gsrs.ncats.nih.gov/) REST API. Retrieve substance metadata, synonyms, cross-reference codes, and chemical structure information for any of the 170 000+ registered substances.

No API key required.

## Installation

```{r}
# install.packages("pak")
pak::pak("c1au6i0/rgsrs")
```

## Functions

| Function | Description |
|---|---|
| `gsrs_search()` | Free-text or Lucene-syntax search |
| `gsrs_substance()` | Substance metadata by UNII (vectorised) |
| `gsrs_names()` | All registered synonyms for a UNII |
| `gsrs_codes()` | Cross-reference codes for a UNII, with optional `code_system` filter |
| `gsrs_unii_from_name()` | Resolve a substance name to its UNII |
| `gsrs_structure()` | Chemical structure data (SMILES, formula, MW, InChI, …) by UNII |
| `gsrs_structure_search()` | Substructure / similarity / exact search by SMILES |
| `gsrs_chem_info()` | Chemical structure info from any identifier: name, CAS, UNII, InChIKey, or SMILES |
| `gsrs_hierarchy()` | Parent/child relationship tree for a UNII |
| `gsrs_all()` | Convenience wrapper: substance + names + codes + structure + hierarchy |
| `gsrs_browse()` | Page through the full GSRS substance catalogue |
| `gsrs_vocabularies()` | Retrieve controlled vocabulary terms |
| `write_dataframes_to_excel()` | Write a named list of data frames to `.xlsx` |

## Usage

```{r}
library(rgsrs)

# Search
gsrs_search("aspirin", top = 5)

# Fetch by UNII (aspirin = R16CO5Y76E)
gsrs_substance("R16CO5Y76E")

# All synonyms
gsrs_names("R16CO5Y76E")

# Cross-references – all systems
gsrs_codes("R16CO5Y76E")

# Cross-references – CAS only
gsrs_codes("R16CO5Y76E", code_system = "CAS")

# Name → UNII
gsrs_unii_from_name(c("aspirin", "ibuprofen"))

# Chemical structure info by name
gsrs_chem_info(c("aspirin", "ibuprofen"), type = "name")

# Chemical structure info by CAS number
gsrs_chem_info(c("50-78-2", "15687-27-1"), type = "cas")

# Chemical structure info by UNII
gsrs_chem_info("R16CO5Y76E", type = "unii")

# Chemical structure info by InChIKey
gsrs_chem_info("BSYNRYMUTXBXSQ-UHFFFAOYSA-N", type = "inchikey")

# Chemical structure info by SMILES (exact match)
gsrs_chem_info("CC(=O)Oc1ccccc1C(=O)O", type = "smiles")

# Everything at once
out <- gsrs_all("R16CO5Y76E")
str(out, max.level = 1)
```

All functions accept a `verbose` argument (default `TRUE`) and return tidy data frames with a `query` column tracking the input identifier. Failed lookups return `NULL` with a warning rather than throwing an error.

## License

MIT © 2026 Claudio Zanettini
