---
output: md_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rgsrs

<!-- badges: start -->
[![R-CMD-check](https://github.com/c1au6i0/rgsrs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/c1au6i0/rgsrs/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`rgsrs` provides a tidy R interface to the
[FDA Global Substance Registration System (GSRS)](https://gsrs.ncats.nih.gov/)
REST API. It allows you to search for substances, retrieve substance metadata
by UNII, look up all registered names (synonyms), fetch cross-reference codes
(CAS, PubChem, ChEMBL, WHO-ATC, etc.), and resolve common names to their UNII.

## Installation

```{r eval=FALSE}
# Install the development version from GitHub:
# install.packages("pak")
pak::pak("c1au6i0/rgsrs")
```

## Core functions

| Function | Description |
|---|---|
| `gsrs_search()` | Free-text or Lucene-syntax search across all substances |
| `gsrs_substance()` | Fetch substance metadata by UNII |
| `gsrs_names()` | All registered synonyms for a UNII |
| `gsrs_codes()` | All cross-reference codes (CAS, PubChem, ChEMBL, …) for a UNII |
| `gsrs_unii_from_name()` | Resolve a substance name to its UNII |
| `gsrs_all()` | Convenience wrapper returning substance + names + codes |
| `write_dataframes_to_excel()` | Write a named list of data frames to Excel |

## Usage

```{r eval=FALSE}
library(rgsrs)

# Search for aspirin
results <- gsrs_search("aspirin", top = 5)

# Fetch substance record by UNII
aspirin <- gsrs_substance("R16CO5Y76E")

# All registered names for aspirin
names_df <- gsrs_names("R16CO5Y76E")

# All external codes (CAS, PubChem, etc.)
codes_df <- gsrs_codes("R16CO5Y76E")

# Only CAS numbers
cas_df <- gsrs_codes("R16CO5Y76E", code_system = "CAS")

# Resolve names to UNIIs
uniis <- gsrs_unii_from_name(c("aspirin", "ibuprofen"))

# Everything at once
all_data <- gsrs_all("R16CO5Y76E")
str(all_data, max.level = 1)
```

## About GSRS

The FDA Global Substance Registration System (GSRS) is a public database
maintained by the National Center for Advancing Translational Sciences (NCATS)
and the FDA.  It assigns a unique UNII (Unique Ingredient Identifier) to every
registered substance and tracks all known names, cross-references, and
structural information.

- API base URL: `https://gsrs.ncats.nih.gov/api/v1`
- No authentication required.
- Rate-limit courtesy: include reasonable delays between bulk requests.

## License

MIT © 2024 Claudio Zanettini
